Analysis of transit delays in the Toronto Transit Commission: 2024

### 2024-09-26 paper/paper.qmd
Update paper.qmd

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(lubridate)

# Define file paths (replace with actual paths)
file1 <- ("ttc_subway_delay_data_2024.csv")
file2 <- ("ttc_bus_delay_data_2024.csv")
file3 <- ("ttc_streetcar_delay_data_2024.csv")

data1 <- read.csv(file1)
data2 <- read.csv(file2)
data3 <- read.csv(file3)

# Inspect column names and structure
print(names(data1))
print(names(data2))
print(names(data3))

# Assuming the date column might be named 'Date'
data1$date <- as.Date(data1$Date, format = "%Y-%m-%d")
data2$date <- as.Date(data2$Date, format = "%Y-%m-%d")
data3$date <- as.Date(data3$Date, format = "%Y-%m-%d")

# Extract the day of the week for each file and count incidents
day_counts_file1 <- data1 %>%
  mutate(day_of_week = wday(date, label = TRUE)) %>%
  group_by(day_of_week) %>%
  summarise(total_count = n()) %>%
  mutate(file = "File 1")

day_counts_file2 <- data2 %>%
  mutate(day_of_week = wday(date, label = TRUE)) %>%
  group_by(day_of_week) %>%
  summarise(total_count = n()) %>%
  mutate(file = "File 2")

day_counts_file3 <- data3 %>%
  mutate(day_of_week = wday(date, label = TRUE)) %>%
  group_by(day_of_week) %>%
  summarise(total_count = n()) %>%
  mutate(file = "File 3")

# Combine the results into a single dataframe
combined_day_counts <- bind_rows(day_counts_file1, day_counts_file2, day_counts_file3)

# Plot the variation of delay incidents by day of the week for each file
ggplot(combined_day_counts, aes(x = day_of_week, y = total_count, color = file, group = file)) +
  geom_line() +
  geom_point() +
  labs(title = "Variation of Delay Incidents by Day of the Week",
       x = "Day of the Week",
       y = "Total Count",
       color = "File") +
  theme_minimal()
```
Figure 3:The trend of delays based on the day of the week

>>>>

```{r}
# Load necessary libraries
library(ggplot2)
library(dplyr)

# Load datasets (replace with actual file paths)
bus_data <- read.csv('ttc_bus_delay_data_2024.csv')
streetcar_data <- read.csv('ttc_streetcar_delay_data_2024.csv')
subway_data <- read.csv('ttc_subway_delay_data_2024.csv')

# Create a mapping of full day names to their abbreviations
day_abbreviations <- c("Monday" = "Mon", "Tuesday" = "Tue", "Wednesday" = "Wed", 
                       "Thursday" = "Thu", "Friday" = "Fri", "Saturday" = "Sat", 
                       "Sunday" = "Sun")

# Function to abbreviate day names
abbreviate_day <- function(day) {
  return(day_abbreviations[day])
}

# Group and filter data for each vehicle type, then abbreviate day names
bus_delays_by_day <- bus_data %>%
  filter(Day %in% names(day_abbreviations)) %>%
  group_by(Day) %>%
  summarise(Count = n()) %>%
  mutate(Day = abbreviate_day(Day))

streetcar_delays_by_day <- streetcar_data %>%
  filter(Day %in% names(day_abbreviations)) %>%
  group_by(Day) %>%
  summarise(Count = n()) %>%
  mutate(Day = abbreviate_day(Day))

subway_delays_by_day <- subway_data %>%
  filter(Day %in% names(day_abbreviations)) %>%
  group_by(Day) %>%
  summarise(Count = n()) %>%
  mutate(Day = abbreviate_day(Day))

# Combine the data
bus_delays_by_day$vehicle <- "Bus"
streetcar_delays_by_day$vehicle <- "Streetcar"
subway_delays_by_day$vehicle <- "Subway"

combined_data <- rbind(bus_delays_by_day, streetcar_delays_by_day, subway_delays_by_day)

# Create the plot with abbreviated day labels
ggplot(combined_data, aes(x = Day, y = Count, color = vehicle, group = vehicle)) +
  geom_line(size = 1) +
  geom_point() +
  labs(title = "Number of delays caused in day, by Vehicle",
       x = "Day of the Week",
       y = "Count of Delays",
       color = "Vehicle") +
  theme_minimal() +
  scale_x_discrete(limits = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))
```
Figure 3:The trend of delays based on the day of the week
